---
title: "Tennis"
output: html_document
---
## Set up environment for analysis
```{r setup, include=FALSE}
library(dplyr)
library(plotly)
library(ggplot2)
#source of files http://www.tennis-data.co.uk/ausopen.php
wta<-read.csv("Data/WTA_ausopen.csv", stringsAsFactors = F)
atp<-read.csv("Data/ATP_ausopen.csv", stringsAsFactors = F)

```

## Clean data
Betting odds were from Bet365 agency
```{r}
# General data cleaning
wta <- wta %>%
  rename(
  tour = WTA)
wta$tour <- "WTA"

atp <- atp %>%
  rename(
    tour = ATP)
atp$tour <- "ATP"

wta <- wta %>% 
  select(tour, Tournament, Date, Round, Winner, Loser, WRank, LRank, WPts, LPts, Comment, B365W, B365L) 
atp <- atp %>% 
  select(tour, Tournament, Date, Round, Winner, Loser, WRank, LRank, WPts, LPts, Comment, B365W, B365L) 

aus_open<-rbind(atp, wta)
aus_open$WRank <- as.integer(aus_open$WRank)
aus_open$LRank <- as.integer(aus_open$LRank)
aus_open$WPts <- as.integer(aus_open$WPts)
aus_open$LPts <- as.integer(aus_open$LPts)

aus_open$Date <- as.Date(aus_open$Date, "%d/%m/%y")

aus_open <- aus_open %>% 
  mutate(favorite = ifelse(B365W > B365L, 0, 1))%>%
  mutate(rank_win = ifelse(WRank > LRank, 0, 1))
aus_open$favorite <- as.factor(aus_open$favorit)
aus_open$rank_win <- as.factor(aus_open$rank_win)
aus_open$year = as.character(format(aus_open$Date, format = "%Y"))
#randomly assign 10% of people as ill
aus_open$W_ill <- as.character(rbinom(n=2794, size=1, prob=0.05))
aus_open$L_ill <- as.character(rbinom(n=2794, size=1, prob=0.05))
```

## Including Plots
### Plot the proportion of correct times the favorite won in Australian Open from 2010 - 2020.
Blue (favorite = 1) represents that the favorite won, while red (favorite = 0) represents the favorite lost
```{r pressure, echo=FALSE}
plotly_build(aus_open%>%
              ggplot(aes(x = year, fill = favorite))+
              geom_bar(aes(y = (..count..)), position = "fill")+
              ggtitle("Proportion of rankings favorites in Aus Open")+
              ylab("proportion"))
```

### Plot the proportion of times the favorite won during from the 4th round.
Blue (favorite = 1) represents that the favorite won, while red (favorite = 0) represents the favorite lost
```{r}
plotly_build(aus_open%>%
              filter(Round == "4th Round"|Round == "Quarterfinals"| Round == "Semifinals" | Round == "The Final")%>%
              ggplot(aes(x = year, fill = favorite))+
              geom_bar(aes(y = (..count..)), position = "fill")+
              ggtitle("Proportion of rankings favorites in Aus Open: Finals")+
              ylab("proportion"))
```

### Plot the proportion of favorites who still won while being sick
Blue (favorite = 1) represents that the favorite won, while red (favorite = 0) represents the favorite lost
```{r}
plotly_build(aus_open%>%
               filter(W_ill == 1) %>% 
               ggplot(aes(x = year, fill = favorite))+
               geom_bar(aes(y = (..count..)), position = "fill")+
               ggtitle("Proportion of winners by betting odds favorites in Aus Open: when winner is sick")+
               ylab("proportion"))
```

### Plot the proportion of favorites that win when the loser is sick.
Blue (favorite = 1) represents that the favorite won, while red (favorite = 0) represents the favorite lost
```{r}
plotly_build(aus_open%>%
               filter(L_ill == 1) %>% 
               ggplot(aes(x = year, fill = favorite))+
               geom_bar(aes(y = (..count..)), position = "fill")+
               ggtitle("Proportion of rankings favorites in Aus Open: when loser is sick")+
               ylab("proportion"))
```

### Look at how often when the non-favorite won the proportion of times the non-favorite was sick.
Blue (L_ill = 1) represents that the favorite was sick, while red (L_ill = 0) represents the favorite wasn't sick.

```{r}
plotly_build(aus_open%>%
               filter(favorite == 0) %>% #filter favorite that lost
               ggplot(aes(x = year, fill = L_ill))+
               geom_bar(aes(y = (..count..)), position = "fill")+
               ggtitle("Proportion of favorites who lost in Aus Open when sick")+
               ylab("proportion"))
```
